---
title: "Sweetpotato Growth Analysis"
output:
  html_document:
    df_print: paged
  keep_tex: yes
  pdf_document: null
---

```{r global_options, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(out.width = "100%", fig.height = 7, fig.path = 'figures/', echo = TRUE, warning = FALSE, message = FALSE)
#fig.with = 25, fig.height = 7
library(googlesheets4)
library(dplyr)
library(tibble)
library("ggpubr")
library(lme4)        # Our primary analysis routine
library(lmerTest)    # A user friendly interface to lme4 that produces p-values
library(emmeans)
library(r2glmm)
library(ggplot2)
library(lemon)
library(stringr)
library(car)
library(forcats)
library(ggrepel)

planting_date <-'2021-06-08'
harvest_date <-'2021-10-01'
```

```{r functions}
fahrenheit_to_celsius <- function(temp_F) {
  temp_C <- (temp_F - 32) * 5 / 9
  return(temp_C)
}

GDD <- function(tmax,tmin){
  t_base = 16.9 #62.42 F (for total biomass)
  t_ceiling = 29.2 #84.56F (for total biomass)
  Tmax <- tmax
  if (tmax > t_ceiling){
    Tmax <- t_ceiling
  }
  if (tmin < 0){
    Tmax <- t_base
  }
  GDD = Tmax - t_base
  return(GDD)
    
}
```

* Importing weights, counts and heights of the sampled biomass components
```{r biomass_import}
weights <- read_sheet("https://docs.google.com/spreadsheets/d/1zYidmi7SMrktiGQ4scQ6V9Ea2xmI3GfMQWD9iOMb_IU/edit?usp=sharing",sheet='biomass_all',na="NaN")

counts <- read_sheet("https://docs.google.com/spreadsheets/d/1efanenzcOrfnaCvHx2MMPASRtx_ZungbIKimgK_-63M/edit?usp=sharing",sheet="Sheet1",na="NaN","Sheet1!A1:H337")

all_biomass = merge(weights, counts, by.x=c("Date","Block","Variety","Plant"), by.y=c("Date","Block","Variety","Plant"))


all_biomass <- all_biomass %>% 
  group_by(Date,Block,Variety) %>% 
  mutate_if(is.numeric,function(x) ifelse(is.na(x),median(x, na.rm = TRUE),x))
all_biomass$Variety <- gsub('Lime', 'Covington',
           gsub('Pink', 'Bellevue',
           gsub('White', 'Monaco', 
          gsub('Blue','Bayou Belle', all_biomass$Variety))))
DAT<-as.numeric(round(all_biomass$Date-as.POSIXct(planting_date)))
all_biomass<-add_column(all_biomass, DAT, .after = 4)
all_biomass$Block <- as.character(all_biomass$Block)
all_biomass$Date <- as.Date(all_biomass$Date)
#all_biomass

```

```{r derived_vars}
all_biomass$Water_content_below<- (all_biomass$Fresh_below-all_biomass$Dry_below)/all_biomass$Dry_below

all_biomass$Water_content_above<- (all_biomass$Fresh_above-all_biomass$Dry_above)/all_biomass$Dry_above

all_biomass$Root_Shoot_ratio<- all_biomass$Dry_below/all_biomass$Dry_above

var_biom <- names(all_biomass)[6:length(names(all_biomass))]
```

```{r all_biomass_summary}
# all_biomass_x <- all_biomass[,c(1,2,3,6:ncol(all_biomass))] %>% group_by(Date,Variety,Block) %>% summarize_if(is.numeric, sum)
ngr<-12
```

* Importing weather data from Kinston weather station and calculating some derived variables like GDDcum
```{r weather_imports}
weather <-read_sheet("https://docs.google.com/spreadsheets/d/1Nrh24GLL0wK2ujy35FxgHxf-AsFk2yAy79U-6qmCvCk/edit?usp=sharing")
weather<-weather[,c('Date','Daily Maximum Air Temperature  (2m) (F)','Daily Minimum Air Temperature (2m) (F)','Average Photosynthetically Active Radiation (2m) (micromol/m2s)')]
names(weather)<-c('Date','tmax','tmin','PAR')
weather<-weather[weather$Date >= planting_date & weather$Date <= harvest_date, ]
weather$Date<-as.Date(weather$Date,tryFormats = c("%Y-%m-%d"))
weather[,c('tmax','tmin')] <- round(fahrenheit_to_celsius(weather[,c('tmax','tmin')]),2)

weather$GDD_daily <- round(mapply(GDD,weather$tmax,weather$tmin),2)
weather$GDD_cum <- cumsum(weather$GDD_daily)

weather$PAR <- round((weather$PAR*0.0864/4.6),2)
weather$PAR_cum <- cumsum(weather$PAR)
#weather
```
* Filtering weather data at each sampling date
```{r weather_at_sampling}
sampling_dates <- unique(all_biomass$Date)
weather_at_sampling <- weather %>% filter(weather$Date %in% sampling_dates)
all_data <-merge(x=all_biomass,y=weather_at_sampling,by="Date",all.x =TRUE)
```

* Setting parameters for figures
```{r plot_pars}
myColors <-c('#225ea8','#f768a1','#fec44f','#bdbdbd')
names(myColors) <- levels(all_biomass$Variety)
custom_colors <- scale_colour_manual(values = myColors)

print("List of variables to analyze")
lista <- var_biom
print(lista)
```


* Inspecting plants variability per block and variety
```{r comparison_all}
# df3=data.frame(newvar = rep(1,7),    
#                Date = as.Date(sampling_dates),
#                newvalue = c(155,178,155,800,800,800,800))

for (var in lista){
  print(paste("Variable: ",var))
  all_biomass0 <- all_biomass
  all_biomass0$variable<-all_biomass0[[var]]
  gpp0<-ggplot(all_biomass0, aes(y=variable, x= Block,color=Variety)) + 
    stat_summary(fun = "mean", geom = "line", 
         alpha = 1,position=position_dodge(0.1)) +
    stat_summary(fun = "mean", geom = "point",
         size = 1.2,position=position_dodge(0.1)) +
    stat_summary(fun.data = "mean_se",
         geom = "errorbar",
         width = .3,position=position_dodge(0.1))+
    custom_colors+ theme_bw()+labs(y = "")+
    #geom_point(data=df3,aes(x=newvar,y=newvalue),colour = "white",alpha=0)+
    facet_wrap(~as.Date(Date), nrow=3,scales="free_y")+labs(y = var)
  gpp<-reposition_legend(gpp0, 'top', panel=c('panel-3-2','panel-3-3'))
}

```

* Repeated linear mixed model
```{r repeated_measurements}
# https://bookdown.org/dereksonderegger/571/10-mixed-effects-models.html
pairwise_rep <- data.frame(Variety=c("Bayou Belle","Bellevue","Covington","Monaco"))
for (var in lista){
  print(paste("Variable:",var))
  all_data0 <- all_data
  all_data0$variable<-all_data0[[var]]
  mix_rep <- lmer(variable ~ Variety+GDD_cum + (1| Block), data=all_data0)
  print("Model summary")  
  summary(mix_rep)
  print("ANOVA")
  print(anova(mix_rep))
  print("Random effects")
  print(ranova(mix_rep))
  r2_rep = r2beta(mix_rep, method = 'kr', partial = T)
  print(r2_rep)
  LetterGroups <- emmeans::emmeans(mix_rep, pairwise~ Variety) %>%
        multcomp::cld(Letters=letters)
  LetterGroups<-LetterGroups[order(LetterGroups$Variety),]
  pairs<-LetterGroups[,c("Variety",".group")]
  pairwise_rep<-merge(pairwise_rep,pairs,by="Variety")
}
names(pairwise_rep)<-c("Variety",lista)
print("All pairwise comparisons")
print(pairwise_rep)

```  

* Mixel linear modeling and pairwise comparison every sampling date
```{r individual_sampling}
pairwise_all <- weather_at_sampling[,c("Date","GDD_cum")][rep(1:7, 4),]
pairwise_all <- pairwise_all[order(pairwise_all$Date),]
pairwise_all$Variety<-rep(unique(all_biomass$Variety),7)
print("Mixel Linear Model and Pairwise comparison")
for (var in lista){
  print(paste("Variable: ",var))
  all_biomass0 <- all_biomass
  all_biomass0$variable<-all_biomass0[[var]]
  LetterResults_all <- NULL
  for (i in 1:7){
    sub_data <- all_biomass0[all_biomass0$Date==sampling_dates[i],]
    nan_count <- sum(is.na(sub_data$variable))
    if (nan_count < 4){
      mixed_var <- lmer(variable ~ Variety + (1|Block), data=sub_data)
      print(Anova(mixed_var,type=c("III")))
      print(ranova(mixed_var))
      r2_sp = r2beta(mixed_var, method = 'kr', partial = T)
      print(r2_sp)
      LetterResults0 <- emmeans::emmeans(mixed_var, pairwise~ Variety) %>%
        multcomp::cld(Letters=letters)
  
      LetterResults <- LetterResults0 %>% 
        mutate(Date = sampling_dates[i], LetterHeight=-10,  .group = str_trim(.group))
      LetterResults_all <- rbind(LetterResults_all, LetterResults[,c("Date","Variety",".group")])
    }
  }
  
  pairwise_all<-merge(pairwise_all,LetterResults_all,by=c("Date","Variety"),all.x=TRUE)

  all_biomass00 <- all_biomass0 %>%
    mutate(Variety = fct_reorder(Variety, variable))
  gg <- ggplot(all_biomass00, aes(y=variable, x= Variety,color=Block)) + 
    facet_wrap(~as.Date(Date), nrow=3,scales="free_y")+
    stat_summary(fun = "mean", geom = "point",
         size = 1.5,position=position_dodge(0.1)) +
    theme_bw()+labs(y = gsub("_", " ", var),x="")+
    geom_text(data=LetterResults_all, aes(label=.group, x= Variety,y=-10),inherit.aes = FALSE)+
    scale_color_brewer(palette = "Set1")+labs(y = var)
  ggp <- reposition_legend(gg, 'top', panel=c('panel-3-2','panel-3-3'))
  ggp
  
}
names(pairwise_all)<-c("Date","Variety","GDD_cum",lista)

```
* Temporal figures and pairwise letters for all growing variables
```{r temporal_plots}
all_biomass_mean <- all_biomass %>% group_by(Date,Variety) %>% 
  summarize_if(is.numeric, mean)
all_biomass_se <- all_biomass %>% group_by(Date,Variety) %>% 
  summarize_if(is.numeric, sd)
all_biomass_se[,5:ncol(all_biomass_se)] <- all_biomass_se[,5:ncol(all_biomass_se)]/sqrt(ngr)

all_mean <- merge(x=all_biomass_mean,y=weather_at_sampling,by="Date",all.x =TRUE)
all_se <- merge(x=all_biomass_se,y=weather_at_sampling,by="Date",all.x =TRUE)


for (var0 in lista){
  print(var0)
  var_mean <- all_mean[,c("GDD_cum","Variety",var0)]
  var_se <- all_se[,c("GDD_cum","Variety",var0)]
  df <-merge(x=var_mean,y=var_se,by=c("GDD_cum","Variety"),all.x=TRUE)
  names(df)<-c("GDD_cum","Variety","var_mean","var_se")
  text_pos <-round(max(df$var_mean+df$var_se,na.rm = TRUE))

  pairs<-pairwise_all[,c("GDD_cum","Variety",var0)]
  df = merge(df,pairs,by=c("GDD_cum","Variety"),all.x=TRUE)
  df$txt <- rep(rep(text_pos,4)+c(0,1,0,1),7)
  names(df)<-c("GDD_cum","Variety","var_mean","var_se","vv","txt")
    
  ggr <- ggplot(df,aes(x=GDD_cum, y= var_mean,color=Variety,label=vv))+geom_line(size = 0.8)+geom_point(size = 2)+
    geom_errorbar(aes(x=GDD_cum, ymin = var_mean-var_se,ymax =var_mean+var_se),size = .4,width=7)+custom_colors+theme_bw()+theme(legend.text=element_text(size=12),legend.position="bottom")+xlab('Accumulated Growing Degree Days (Celsius-based)')+ylab(gsub("_", " ", var0))+scale_x_continuous(breaks = round(unique(df$GDD_cum)))+ geom_text_repel(aes(y=txt),segment.color = NA,size = 4,box.padding=0.05,show.legend = FALSE)
  print(ggr)
}

```

```{r}



```

```{r}


```